---
title: "tidytides"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidaltools)
```

## Goal

This `R` package contains tools to work with the Portuguese National Hydrographic Institute's tidal height data - the *tidal tables*.

It has 2 main tasks:

1. Downloading the tidal tables for chosen time periods
2. Estimate tidal height at given time points


Task 1 is accomplished via their REST API. 

Task 2 takes the heigts from the tidal tables and interpolates it to the chosen time points by approximating it as a sin-wave ([reference in Portuguese](https://www.hidrografico.pt/recursos/tabmares/2021/TabelaMares_Capitulo1_Generalidades_2021.pdf)).


This package started a side-project and was meant to be more of a personal tool than a public one. Since then it has evolved a lot, but assumption checks, error handling and error reporting are still not great. If you use this tool, make sure you read how data should be passed into the arguments. The package has standard R formatted documentation, meaning you can always use the `?` operator to see the help for a function. 


## How to use it?

### Select region

To find the ID of the closest port to your reference area use `port_list()` to see a list of the ports with available tidal tables. This is a list that I compiled myself. If you think that ID is wrong, you can use `get_port_list()` to query a new list of IDs (this will take a while and should be avoided)

### Download tidal tables

To download the tidal tables you can use `get_tides`, which will return a data.frame with the tidal events for the chosen port and dates. Check `?get_tides` to see how to use it. Times are returned given both in local (with daylight savings) and in UTC time.

```{r get_tides}
# Get tidal tables for Faro-Olhão, starting 5 March 2020 and over 2 days
tidal_table <- get_tides(port_id = 19, date = "2020-03-05", day_range = 2)

tidal_table
```

### Interpolate tides

You can use `interpolate_tides` to estimate tidal heights at your chosen port and time points. You also specify if your time points are in "local" or "UTC" time. 

```{r interpolate_tides}
# Times for which I want estimted tidal heights
sampling_times <- c("2020-03-13 15:15:00", "2020-03-09 16:15:00")

# Estimate tidal heights for Faro-Olhão harbor. The times are in local time
interpolate_tides(date_times = sampling_times, port_id = 19, timezone = "local")
```



