---
title: "tidytides"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidytides)
```

## Goal

This is an R package to read and process tidal data from the Portuguese National Hydrographic Institute.

With it, you can retrieve tidal tables programmatically, estimate tidal height at desired points in time and create simple calendar plots of tidal events to help you plan field work.

Initially, this package was built as a quick-fix for some work I had to do, and was more of a personal tool than a public one. As such, assumption checks, error handling and error reporting are not very sophisticated (although this might be more attributed to my ignorance on these subjects). 

If you use this tool, make sure you read how data should be passed into the arguments. The package has standard R formatted documentation, meaning you can always use the `?` operator to see the help for a function. 


## Functionality

### 1. See port IDs

To retrieve a tidal table, you must specify what port you want. This package includes 2 ways to do that: 

#### *port_list* 

Returns a pre-compiled data.frame with available ports. These ids were retrieved in February 2020, if you are getting errors with the IDs in this table, see *get_port_list*.

**usage:** 

port_list()


#### *get_port_list*

**DO NOT USE UNLESS THE IDS FROM port_list() FAIL** Brute forces a test for all IDs from 0 to 1000 and returns all valid ones. While I doubt that this would actually put any significant load on the servers, brute force methods are a little rude so this function is only a back up in case the pre-compiled IDs fail. It also takes some time to run so it's not as convenient.

**usage:** 

get_port_list()

### 2. Retrieve a tidal table

#### *get_tides*

Returns a data.frame with all tidal events for a time period.

port_id - Integer - The id code for the desired port. Use `port_list()` to see a list of IDs. Defaults to Faro-Olhão.
date - Character or POSIXct - The starting date for the wanted tides. Format should be yyyy-mm-dd or yyyy/mm/dd. Defaults to current date
day_range - Integer - The number of days after date for which to retrieve information
include_moons - Logical - Should lunar events be kept in the table? Defaults to FALSE

**usage:** 

```{r get_tides}
table <- get_tides(port_id = 19, date = "2020-03-05", day_range = 7, include_moons = FALSE)

table
```

### 
#### *interpolate_tides*

**usage** sampling_times <- c("2020-03-13 15:15:00", "2020-03-09 16:15:00")
interpolate_tides(date_times = sampling_times, port_id = 19)



Returns a vector with estimated tidal height, at the time points supplied in date_times.
date_times A character vector with the format yyyy-mm-dd hh:mm:ss or POSIXct vector
port_id The id code for the desired port (use port_list to see a list, Faro-Olhão is the default)

#### *plot_tides*

**usage:** plot(tide_table)

Takes a table returned by **get_tides** and turns it into a calendar-style plot. Ideal for quickly looking at when tidal events occur over a time period. I recomment adding a geom_vline at times of interest to make interpretation easier and quicker. Used, mostly, to plan field work.
